---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(opendatasus)
```

# opendatasus

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/danielvartan/opendatasus/workflows/R-CMD-check/badge.svg)](https://github.com/danielvartan/opendatasus/actions)
[![Codecov test coverage](https://codecov.io/gh/danielvartan/opendatasus/branch/main/graph/badge.svg)](https://app.codecov.io/gh/danielvartan/opendatasus?branch=main)
[![License: MIT](https://img.shields.io/badge/license-MIT-green)](https://choosealicense.com/licenses/mit/)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-v2.0%20adopted-ff69b4.svg)](https://danielvartan.github.io/opendatasus/CODE_OF_CONDUCT.html)
<!-- badges: end -->

## Overview

`opendatasus` is an user-friendly R package designed for downloading and processing data from the [DATASUS](https://datasus.saude.gov.br/)'s [openDataSUS platform](https://opendatasus.saude.gov.br).

## Prerequisites

To use the main functions of `opendatasus`, some familiarity with the [R programming language](https://www.r-project.org/) is required.

If you are not comfortable with R, we strongly recommend checking out Hadley Wickham and Garrett Grolemund's free online book [R for Data Science](https://r4ds.hadley.nz/), as well as the Coursera course [Data Science: Foundations using R](https://www.coursera.org/specializations/data-science-foundations-r) from John Hopkins University.

## Installation

You can install `opendatasus` using:

``` {r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("danielvartan/opendatasus")
```

## Usage

We're working on new API functions to make it easier to access the data available on the openDataSUS platform. You can find a list of available functions [here](https://danielvartan.github.io/opendatasus/reference/index.html).

Example of use:

```{r}
library(opendatasus)

fetch_demas_sisvan(uf = "SP") |>
  dplyr::glimpse()
```

## Citation

If you use `opendatasus` in your research, please consider citing it. We have put a lot of work into building and maintaining this free and open-source R package. The citation can be found below.

```{r}
citation("opendatasus")
```

## Contributing

We welcome contributions, including bug reports.

Please take a moment to review our [Guidelines for Contributing](https://danielvartan.github.io/opendatasus/CONTRIBUTING.html).

<br>

Become a supporter of `opendatasus`!

Click [here](https://github.com/sponsors/danielvartan) to make a donation. Please indicate the `opendatasus` package in your donation message.
